---
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: false
  number_sections: true
 
title: "Adding ISBNs to Wiki books"
author: "[Matthew-D-Dwyer](https://github.com/Matthew-D-Dwyer)"
date: "`r paste0('Last Run: ', format(Sys.time(), '%A %d-%B-%Y'))`"
params: 
  param1: "Don't Forget about params"

---

<style>

#TOC {
 font-family: Calibri; 
 font-size: 16px;
 border-color: #3D68DF;
 background: #3D68DF;
}

body {
  font-family: Garamond;
  font-size: 16px; 
  border-color: #D0D0D0;
  background-color: #D0D0D0;
  color: #1A1A1A;
}

pre {
  color: #1A1A1A
  background: #D0D0D0;
  background-color: #D0D0D0
  font-family: Calibri; 
  
}

</style>

```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = TRUE)

knitr::opts_chunk$set(collapse = TRUE)

knitr::opts_chunk$set(warning = TRUE)

knitr::opts_chunk$set(message = TRUE)

knitr::opts_chunk$set(include = TRUE)

custom_black <- '1A1A1A'
custom_white <- 'C0C0C0'
custom_grey_dark <- '6F6F6F'
custom_grey_light <- 'B2B2B2'
custom_accent_blue <- '3D6BFF'

```

```{r libraries etc., message = FALSE }

library(tidyverse)
library(dplyr)
library(RSQLite)

```

### How many books? 

The first thing I wanted to check is how many books there are in each of the different languages.

This code threw errors on my machine indicating that not all rows were retrieved however I was able to check the source and my result is a full data set. 

```{r dataload}

wb_data_base <- file.path('/Users',
                          'macatron',
                          'Documents',
                          'Data Analysis',
                          'R',
                          'wiki--books',
                          'wikibooks.sqlite')

#creating connection
con <- DBI::dbConnect(RSQLite::SQLite(), wb_data_base)

# Lookup to get the full language from the two character part 
Languages <- data.frame(Abbv = c('de',
                                 'en',
                                 'es',
                                 'fr',
                                 'he',
                                 'hu',
                                 'it',
                                 'ja',
                                 'nl',
                                 'pl',
                                 'pt',
                                 'ru'), 
                        full_name = c('German', 
                                      'English', 
                                      'Spanish', 
                                      'French', 
                                      'Greek',
                                      'Hungarian',
                                      'Italian',
                                      'Japanese',
                                      'Dutch',
                                      'Polish',
                                      'Portuguese',
                                      'Russian'))

names_stripper <- function(full_title) 
  #Function that removes 'Wikibooks: ' from the start of the title and everything after and including the first slash
{
  title_no_prefix <- full_title %>%
  str_remove_all('Wikibooks: ')
  
  slash_loc <- str_locate(title_no_prefix, '/')[1]
  
  if (is.na(slash_loc)) {
    title_no_suffix <- title_no_prefix } 
  
  else if (slash_loc > 0) {
  title_no_suffix <- title_no_prefix %>%
                     str_sub(1, slash_loc-1)} else {
      print('Error Error Does Not Compute')
      title_no_suffix <- 'Error Error Does Not Compute'
    }
  
  title_no_suffix }

title_counter <- function(languages_brv) 
  # function that counts how many titles there ar in the data base for that language 
{
  
  con <-  DBI::dbConnect(RSQLite::SQLite(), wb_data_base)   
  
  x <- dbSendQuery(con, paste0("SELECT count (*) FROM ", languages_brv), n = Inf) %>%
    dbFetch(n =Inf)
  
  x <- data.frame(titles = x, Abbv = languages_brv)
  
}
  
Titles_table <- map_dfr(Languages$Abbv, title_counter)

Titles_table <- Titles_table %>%
  left_join(Languages, by = c('Abbv' = 'Abbv')) %>%
  arrange(desc(count....)) 

Titles_table <- Titles_table %>%
   mutate(Language = factor(full_name, levels = Titles_table$full_name)) %>%
  rename(Titles = count....)

```

### Plotting the number of books in each language

The vast majority of books are English 

```{r }

Titles_table %>%
  ggplot(aes(x = Language, y = Titles)) +
  geom_col(fill = 'cornflower blue') +
  geom_label(aes(label = Titles), size = 2.5) +
  theme_minimal() +
  labs(title = 'Wikibooks Number of Titles by Language', 
       subtitle = 'Extraction date: Sep 09 2021',   
       caption = 'Source: Kaggle, starter-wikibooks-dataset') +
  coord_flip() 

```

### Trying to extract the ISBN 

The book title information and the data base in general is pretty messy. 
But the full text of the books often contains the ISBN and if this number can be extracted, even for a fraction of the books then this can be used to link the full text to higher quality metadata. 

Starting with one book that does have ISBN in it. 

```{r}

#Pulling the first English Book
Sample_full_text <- dbSendQuery(con, "SELECT * FROM en", 
                                 n = 1) %>% 
                     dbFetch(n = 1)

ISBN_Loc <- str_locate(Sample_full_text$body_text, "ISBN: ")

```

Looks like no ISBN in that one, but what if I test it for some text I know is 
in the book.

```{r}

Oncology_loc <- str_locate(Sample_full_text$body_text, "Oncology")

```

Test passed so it won't work for that book, what if I cycle through 1000 books?

Again this throws errors but the result has 1,000 observations so I'm OK. 

```{r}

books_1000 <- dbSendQuery(con, "SELECT * FROM en", 
                                 n = 1000) %>% 
                     dbFetch(n = 1000)

```

Cycling through them all, printing found one if the 'ISBN: ' search returns a result

```{r}

book_list <- books_1000$body_text 

for (book in book_list) {
  
  ISBN_Loc <- str_locate(book, 'ISBN: ')
  
  if (!is.na(ISBN_Loc[1])) {
    print('Found One!')
    
  }}

```

Looks like there was one in that first thousand, does that mean about 83 in the whole English data set? 

Again with the errors, but all the books load.

```{r}

all_books <- dbSendQuery(con, "SELECT * FROM en", 
                                 n = Inf) %>% 
                     dbFetch(n = Inf)

Books_with_ISBN <- data.frame(title = NULL, 
                              ISBN_Loc = NULL,
                              Book_no = NULL)

for (i in 1:length(all_books$title)) {
  
  ISBN_Loc <- str_locate(all_books$body_text[i], 'ISBN')
  
   if (!is.na(ISBN_Loc[1])) {
     
     Books_with_ISBN <- rbind(Books_with_ISBN, 
                             data.frame(title = all_books$title[i], 
                                        ISBN_Loc = ISBN_Loc[1],
                             Book_no = i))
     
   } }

print('Books with a possible ISBN: ')

print(length(Books_with_ISBN$title))

```

Initially I only got 84 which is exactly 1 in a thousand, but then I re ran it without the colon after ISBN this got a much higher (but still lower than I hoped) hit rate of about 30% (2481 books). 

### Pulling the actual ISBNs

An ISBN is 10 or 13 digits not including the hyphens colons spaces or whatever

So to start I'll just take the next 30 characters - to be cleaned up later.

Printing 6 at the end so we can get an idea of what further cleaning may be required.

```{r}

indexes <- Books_with_ISBN$Book_no 

ISBN_Loc <- Books_with_ISBN$ISBN_Loc

isbn <- c()

Books_with_ISBN$Full_Text <- all_books$body_text[indexes]

Books_with_ISBN$ISBN <- substr(Books_with_ISBN$Full_Text, 
                               Books_with_ISBN$ISBN_Loc,
                               Books_with_ISBN$ISBN_Loc + 30)
# Removing the full text to make it a bit more manageable copute wise

Books_with_ISBN <- Books_with_ISBN %>% select(-Full_Text)

head(Books_with_ISBN$ISBN)

```

Looking like a good start, all these records above look like they can be cleanable into a real ISBN numbers. 

### Cleaning the ISBNs 

First remove anything that isn't a number and make into a single character. 

```{r}

isbn_cleaner <- function(un_clean_isbn) {
  
  un_clean_isbn %>% 
                      str_extract_all("[0-9]") %>%
                      unlist() %>% 
                      str_c(collapse = '') }

Books_with_ISBN$ISBN_clean <- map_chr(Books_with_ISBN$ISBN, 
                                     isbn_cleaner)

Books_with_ISBN$ISBN_clean[1:20]

```

Looking at the above first 20 ISBN's it it looking pretty good. 

ISBNs should have 10 or 13 digits lets see what the distribution is like. 

### Extracting German ISBNs

```{r }

german_books <-  dbSendQuery(con, "SELECT * FROM de") %>% 
                     dbFetch()

book_list <- german_books$body_text 

for (book in book_list) {
  
  ISBN_Loc <- str_locate(book, 'ISBN')
  
  if (!is.na(ISBN_Loc[1])) {
    print('Found One!')
    
  }}

```

Looks like lots of German ones have them!

Making a table of them 

```{r}
### Note need to change for German books
indexes <- Books_with_ISBN$Book_no 

ISBN_Loc <- Books_with_ISBN$ISBN_Loc

isbn <- c()

Books_with_ISBN$Full_Text <- all_books$body_text[indexes]

Books_with_ISBN$ISBN <- substr(Books_with_ISBN$Full_Text, 
                               Books_with_ISBN$ISBN_Loc,
                               Books_with_ISBN$ISBN_Loc + 30)
# Removing the full text to make it a bit more manageable copute wise

Books_with_ISBN <- Books_with_ISBN %>% select(-Full_Text)

head(Books_with_ISBN$ISBN)
```


What about Japanese? Will the same method work



### Testing method on Japanese title

```{r}

japanese_books <-  dbSendQuery(con, "SELECT * FROM ja") %>% 
                     dbFetch()

book_list <- japanese_books$body_text 

for (book in book_list) {
  
  ISBN_Loc <- str_locate(book, 'ISBN')
  
  if (!is.na(ISBN_Loc[1])) {
    print('Found One!')
    
  }}

```

Making a table of them, how screwy are they? 

```{r }

### Note need to change for German books
indexes <- Books_with_ISBN$Book_no 

ISBN_Loc <- Books_with_ISBN$ISBN_Loc

isbn <- c()

Books_with_ISBN$Full_Text <- all_books$body_text[indexes]

Books_with_ISBN$ISBN <- substr(Books_with_ISBN$Full_Text, 
                               Books_with_ISBN$ISBN_Loc,
                               Books_with_ISBN$ISBN_Loc + 30)
# Removing the full text to make it a bit more manageable copute wise

Books_with_ISBN <- Books_with_ISBN %>% select(-Full_Text)

head(Books_with_ISBN$ISBN)

```
